Higher-Order unification for free!

### 2024-04-09 paper/main.tex
small changes

% This property can be relaxed, to accept approximations: a dedicated
% section is given in a future section
\begin{lemma}[\linkbeta with rigid \lhs]
  If the \lhs of a \linkbeta is instantiated to a rigid term and its \rhs
  counterpart is still in \notllambda,
  the original unification problem is not in \llambda
  % \noteE[inline]{non capisco cosa vuoi dire con "the current...". Data la prova sotto vuoi dire che il problem originale sarebbe stato fuori... e quindi falliamo}
  %\noteE[inline]{Is it clearer?}
  and the unification fails.
  \label{lemma:keep-fidelity}
\end{lemma}

\begin{proof}[Proof sketch]
  Given $X \appsep t_1\dots t_n \Ue t$ where $t$ is a rigid term and
  $t_1\dots t_n$ is not in \llambda. By construction, $X \appsep t_1\dots t_n$
  is replaced with a variable $Y$, and the \linkbeta \linkbetaM{\Gamma}{Y}{X
  \appsep t_1\dots t_n} is created. If $Y$ is instantiated to the rigid term $t$,
  and \rhs is still in \notllambda, then the original \Fo{} unification problem
  is between two terms in \notllambda. Unification fails.
\end{proof}

@@

\newcommand{\progBetaLL}{\emph{progress-beta-\llambda}\xspace}
\newcommand{\progBetaRH}{\emph{progress-beta-rigid-head}\xspace}
\newcommand{\progBetaDedup}{\emph{progress-beta-dedup}\xspace}

@@

\noteD{added the def below to respect fidelity}
% Example of this `X a = Y a'
\begin{definition}[\progBetaDedup]
  Given a \linkbeta $l_1$ and second link $l_2 \in\linkStore$, such that they
  share the same \lhs. In this case, the two \rhs are unified and a $l_2$ is
  removed from \linkStore.
  \label{def:progBetaDedup}
\end{definition}

@@

\begin{proof}[Proof sketch]
  Let $l$ a \linkbeta in the store \linkStore. If $l$ is activated by
  \progBetaRH, then it disappears from \linkStore and \elpiIn{progress}
  terminates. Otherwise, the \rhs of $l$ is made by a variable applied to $m$
  arguments. At each activation of \progBetaLL, $l$ is replaced by a new
  \linkbeta $l^1$ having $m-1$ arguments. At the $m^{th}$ iteration, the
  \linkbeta $l^m$ has no more arguments and is removed from \linkStore.
  Note that at the $m^{th}$ iteration, $m$ new \linketa have been added to
  \linkStore, however, by \cref{lemma:prog-eta-terminates}, the algorithm
  terminates. Finally \progBetaDedup also guarantees termination since it 
  makes a unification $\mathbb{U}$ and if $\mathbb{U}$ fails, then 
  \elpiIn{progress} terminates and if $\mathbb{U}$ succeeds, 
  the recursive calls to \elpiIn{progress} have a specialized \linkbeta\dots

@@

\begin{corollary}
  Given a \linkbeta, the variables occurring in its \rhs are in \llambda.
  % \noteE[inline]{vuoi dire: le variabili che occorrono nel rhs sono in...?}
  \noteD[inline]{is it clearer?}
\end{corollary}

\begin{proof}[Proof sketch]
  By construction, the \rhs of \linkbeta has the shape \rhsBeta, $s_1\dots s_n$
  is in \llambda and all the terms $t_1\dots t_n$ are in \llambda, too. If a
  \linkbeta is triggered by \progBetaRH, then, by \cref{def:progBetaRH}, that
  link is removed by \linkStore, and the property is satisfied. If the \linketa
  is activated by \progBetaLL, then, by \cref{def:progBetaLL}, the new \linkbeta
  as a variable as a scope which is still in \llambda. 
\end{proof}

\begin{lemma}[Fidelity with \linkbeta]
  The introduction of \linkbeta guarantees \cref{prop:fidelity} 
\end{lemma}

\begin{proof}[Proof sketch]
  Let \hoUnifPb a unification problem and $\sigma$ a substitution
  such that $\hoUnifPb \in \notllambda$. 
  If $\sigma\hoUnifPb$ is in \llambda, then by
  \cref{def:progBetaLL,def:progBetaRH}, the \linkbeta associated to the subterm of
  \hoUnifPb have been solved and removed. The
  unification is done between terms in \llambda and by \cref{lemma:fidelity-maybeeta}
  fidelity is guaranteed. If $\sigma\hoUnifPb$ is in \notllambda,
  then, by \cref{lemma:keep-fidelity}, the unification fails, as per the
  corresponding unification in \Fo{}.
\end{proof}

>>>>

% This property can be relaxed, to accept approximations: a dedicated
% section is given in a future section

@@

\newcommand{\progBetaLL}{\emph{progress-beta-\llambda}\xspace}
\newcommand{\progBetaRH}{\emph{progress-beta-rigid-\rhs}\xspace}
\newcommand{\progBetaDedup}{\emph{progress-beta-dedup}\xspace}
\newcommand{\progBetaLH}{\emph{progress-rigid-\lhs}\xspace}

@@

% Example of this `X a = Y a'
\begin{definition}[\progBetaDedup]
  Given a \linkbeta $l_1$ and second link $l_2 \in\linkStore$, such that they
  share the same \lhs. The two \rhs are unified and a $l_2$ is
  removed from \linkStore.
  \label{def:progBetaDedup}
\end{definition}

\begin{definition}[\progBetaLH]
  Given a \linkbeta with rigid \lhs, the unification fails.
  \label{def:progBetaLH}
\end{definition}

@@

\begin{proof}[Proof sketch]
  Let $l$ a \linkbeta in the store \linkStore. If $l$ is activated by
  \progBetaRH, then it disappears from \linkStore and \elpiIn{progress}
  terminates. Otherwise, the \rhs of $l$ is made by a variable applied to $m$
  arguments. At each activation of \progBetaLL, $l$ is replaced by a new
  \linkbeta $l^1$ having $m-1$ arguments. At the $m^{th}$ iteration, the
  \linkbeta $l^m$ has no more arguments and is removed from \linkStore.
  Note that at the $m^{th}$ iteration, $m$ new \linketa have been added to
  \linkStore, however, by \cref{lemma:prog-eta-terminates}, the algorithm
  terminates. Finally \progBetaDedup (resp. \progBetaLH) also guarantees termination since 
  it removes a link from \linkStore (resp. immediately fails).

@@

% \begin{corollary}
%   Given a \linkbeta, the variables occurring in its \rhs are in \llambda.
%   % \noteE[inline]{vuoi dire: le variabili che occorrono nel rhs sono in...?}
%   \noteD[inline]{is it clearer?}
% \end{corollary}

% \begin{proof}[Proof sketch]
%   By construction, the \rhs of \linkbeta has the shape \rhsBeta, $s_1\dots s_n$
%   is in \llambda and all the terms $t_1\dots t_n$ are in \llambda, too. If a
%   \linkbeta is triggered by \progBetaRH, then, by \cref{def:progBetaRH}, that
%   link is removed by \linkStore, and the property is satisfied. If the \linketa
%   is activated by \progBetaLL, then, by \cref{def:progBetaLL}, the new \linkbeta
%   as a variable as a scope which is still in \llambda. 
% \end{proof}

\begin{lemma}[Fidelity with \linkbeta]
  The introduction of \linkbeta guarantees \fullref{prop:fidelity} 
\end{lemma}

\begin{proof}[Proof sketch]
  Let \hoUnifPb a unification problem and $\sigma$ a substitution
  such that $\hoUnifPb \in \notllambda$. 
  If $\sigma\hoUnifPb$ is in \llambda, then by
  \cref{def:progBetaLL,def:progBetaRH}, the \linkbeta associated to the subterm of
  \hoUnifPb have been solved and removed. The
  unification is done between terms in \llambda and by \cref{lemma:fidelity-maybeeta}
  fidelity is guaranteed. If $\sigma\hoUnifPb$ is in \notllambda,
  then, by \cref{def:progBetaLH}, the unification fails, as per the
  corresponding unification in \Fo{}.
\end{proof}
